### 2019-05-31 Alternatives-for-Neighborhood-Function-in-Kohonen-Maps/presentation.tex
Presentation was added.

\documentclass{beamer}

\mode<presentation> {
	\usetheme{Berlin}
}

\title[12th International Conference on Large-Scale Scientific Computations, June 10 - 14, 2019, Sozopol, Bulgaria]{
	Alternatives for Neighborhood Function in Kohonen Maps
}

\author{Iliyan Zankinski, Kolyu Kolev, Todor Balabanov}

\date{10-14.VI.2019}

\institute[IICT-BAS, LSSC'19] {
	Institute of Information and Communication Technologies \\ 
	Bulgarian Academy of Sciences \\
	\medskip
	\textit{iliyan@hsi.iccs.bas.bg}
}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\frametitle{Overview}
\tableofcontents
\end{frame}

\section{Introduction}

\begin{frame}
\center \huge{Introduction}
\end{frame}

\begin{frame}
\frametitle{Clustering}
\begin{itemize}
  \item Separation of the data set in set of groups
  \item Straight clustering - each data sample belongs exactly to only one group
  \item Fuzzy clustering - each data sample has varying degree of membership to different groups
\end{itemize}
Optimal clustering is data set separation which minimizes the distance inside the cluster and maximizes the distance between the clusters
\end{frame}

\begin{frame}
\frametitle{Self-Organizing Maps}
\begin{itemize}
  \item Proposed by Teuvo Kohonen in the 1980s
  \item Have application in meteorology, oceanography, project prioritization and selection, seismic facies analysis for oil and gas exploration, failure mode and effects analysis, creation of artwork and many other areas
  \item They are very useful for visualization by data dimensions reduction
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{SOM in General}
\begin{itemize}
  \item Unsupervised competitive learning is used
  \item Classify input data in predefined number of clusters
  \item When the net has fewer nodes it achieve results similar to K-means clustering
\end{itemize}
\end{frame}

\section{Neighborhood Function Proposition}

\begin{frame}
\center \huge{Neighborhood Function Proposition}
\end{frame}

\begin{frame}
\frametitle{SOM Training}
\begin{itemize}
  \item The network consists of a grid with units
  \item The units are connected to adjacent units with a neighborhood relation
  \item The network forms an elastic mesh
  \item Data samples lying near each other in the input space are mapped into nearby grid units
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Neighborhood Function}
\begin{itemize}
  \item It gives scaling factor for the distance between one neuron and other neurons in each step
  \item The simplest form of a neighborhood function gives 1 for the closest nodes and 0 for all other
  \item The most used neighborhood function is a Gaussian function
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Fading Cosine}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\linewidth]{fig01a}
\label{fig01a}
\end{figure}
\begin{equation} \label{equ01}
f(x) = 
	\begin{cases} 
		cos(x) & -\pi/2 < x < \pi/2 \\
		\frac{cos(x)}{|x|} & +\pi/2 \geq x \leq -\pi/2 \\
	\end{cases}
\end{equation}
\end{frame}

\begin{frame}
\frametitle{Exponential Regulated Cosine}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\linewidth]{fig01b}
\label{fig01b}
\end{figure}
\begin{equation} \label{equ02}
f(x) = \frac{cos(x)}{e^{|x|}}
\end{equation}
\end{frame}

\section{Experiments and Results}

\begin{frame}
\center \huge{Experiments and Results}
\end{frame}

\begin{frame}
\frametitle{Experimental Setup}
\begin{itemize}
  \item All experiments are done in C as a programming language
  \item Single processor desktop machine - Intel Core i5, 2.3 GHz, 2 Cores, 8GB RAM and Mac OS X 10.13.6, Apple LLVM version 9.1.0
  \item The source code is available at: 
  http://github.com/Coding-Sunday-Sofia/Kohonen
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Experimental Data}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\textwidth]{fig03}
  \label{fig03}
\end{figure}
Rents and living area data set
\end{frame}

\begin{frame}
\frametitle{Gausssian Test}
\begin{figure}[h]
  \centering
  \includegraphics[width=1.0\textwidth,height=0.6\textwidth]{fig02a}
  \label{fig02a}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Fading Cosine Test}
\begin{figure}[h]
  \centering
  \includegraphics[width=1.0\textwidth,height=0.6\textwidth]{fig02b}
  \label{fig02b}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Exponential Regulated Cosine Test}
\begin{figure}[h]
  \centering
  \includegraphics[width=1.0\textwidth,height=0.6\textwidth]{fig02c}
  \label{fig02c}
\end{figure}
\end{frame}

\section{Conclusions}

\begin{frame}
\center \huge{Conclusions}
\end{frame}

\begin{frame}
\frametitle{Concluding Remarks}
\begin{itemize}
  \item The proposed neighbor functions give promising results
  \item Stressing the neighboring nodes down and up leads to better separation between the formed clusters
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Questions and Answers}
\center \huge{Thank you for the attention!}
\end{frame}

\end{document}